### 2023-10-12 emu_JHEP_REsubmission/responseToEditor.txt
added subdirectory, shared EMU.nb notebook

Dear Editor,

We thank the referee for their feedback yada yada yada

We have addressed all the minor clarifications in addition to the following typos we have identified through an a final read through:

C1] 5.46 and 5.47 swapped the generating functions of H(ij)(kl) and H(ijkl) -- this has been resolved.
C2] 5.23 had a misplaced '+' that has been removed
...

Furthermore, we appreciative the substance and less superficial feedback and have made the following additions to the manuscript and provide the following clarification,

Q1]
Q2]
...

Sincerely,
JJMC, NHP